---
title: "Practical Machine Learning - Course Project"
author: "Rodrigo Rivera"
date: "August 28, 2016"
output: html_document
---

# Background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

#Data
The training data for this project are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv
The test data are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv
The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har.

##Load Libraries
```{r}
library(AppliedPredictiveModeling)
library(caret)
library(ElemStatLearn)
library(pgmm)
library(rpart)
library(gbm)
library(lubridate)
library(forecast)
library(e1071)
library(randomForest)
library(rattle)
```

##Download The Data
```{r}
trainUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
train <- read.csv(url(trainUrl), na.strings=c("NA","#DIV/0!",""))
testUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
test <- read.csv(url(testUrl), na.strings=c("NA","#DIV/0!",""))
dateDownloaded <- today()
```

##Creating the Training Data Partition
```{r}
inTrain <- createDataPartition(y=train$classe, p=0.6, list=FALSE)
training <- train[inTrain, ]
testing <- train[-inTrain, ]
dim(training)
dim(testing)
```

##Cleaning The Data Set
Finding and eliminating Near Zero Variance variables
```{r}
DataNZV <- nearZeroVar(training, saveMetrics=TRUE)
NZV <- DataNZV$nzv==T
myNZVvars <- names(myTraining) %in% c("new_window", "kurtosis_roll_belt", "kurtosis_picth_belt",
"kurtosis_yaw_belt", "skewness_roll_belt", "skewness_roll_belt.1", "skewness_yaw_belt",
"max_yaw_belt", "min_yaw_belt", "amplitude_yaw_belt", "avg_roll_arm", "stddev_roll_arm",
"var_roll_arm", "avg_pitch_arm", "stddev_pitch_arm", "var_pitch_arm", "avg_yaw_arm",
"stddev_yaw_arm", "var_yaw_arm", "kurtosis_roll_arm", "kurtosis_picth_arm",
"kurtosis_yaw_arm", "skewness_roll_arm", "skewness_pitch_arm", "skewness_yaw_arm",
"max_roll_arm", "min_roll_arm", "min_pitch_arm", "amplitude_roll_arm", "amplitude_pitch_arm",
"kurtosis_roll_dumbbell", "kurtosis_picth_dumbbell", "kurtosis_yaw_dumbbell", "skewness_roll_dumbbell",
"skewness_pitch_dumbbell", "skewness_yaw_dumbbell", "max_yaw_dumbbell", "min_yaw_dumbbell",
"amplitude_yaw_dumbbell", "kurtosis_roll_forearm", "kurtosis_picth_forearm", "kurtosis_yaw_forearm",
"skewness_roll_forearm", "skewness_pitch_forearm", "skewness_yaw_forearm", "max_roll_forearm",
"max_yaw_forearm", "min_roll_forearm", "min_yaw_forearm", "amplitude_roll_forearm",
"amplitude_yaw_forearm", "avg_roll_forearm", "stddev_roll_forearm", "var_roll_forearm",
"avg_pitch_forearm", "stddev_pitch_forearm", "var_pitch_forearm", "avg_yaw_forearm",
"stddev_yaw_forearm", "var_yaw_forearm")
myTraining <- myTraining[!myNZVvars]
dim(myTraining)
clean1 <- colnames(myTraining)
clean2 <- colnames(myTraining[, -58])
myTesting <- myTesting[clean1]
testing <- testing[clean2]
```

Eliminating first column of the dataset because it may intefere with ML Algorithms
```{r}
myTraining <- myTraining[c(-1)]
```

Eliminating NAs and changing them for the mean
```{r}
preProcess(myTraining, method="knnImpute")
```

#Predictions
##Decision Tree
```{r}
fit1 <- rpart(classe ~ ., data=myTraining, method="class")
fancyRpartPlot(fit1)
predict1 <- predict(fit1, myTesting, type = "class")
confusionMatrix(predict1, myTesting$classe)$overall[1] #accuracy
```

##Random Forest
```{r}
fit2 <- train(classe ~ ., data=myTraining, method="rf")
predict2 <- predict(fit2, myTesting, type = "class")
confusionMatrix(predict2, myTesting$classe)$overall[1] #accuracy
```

##Boosted Trees
```{r}
fit3 <- train(classe ~ ., data=myTraining, method="gbm")
predict3 <- predict(fit3, myTesting, type = "class")
confusionMatrix(predict3, myTesting$classe)$overall[1] #accuracy
```

##Linear Discriminant Analysis
```{r}
fit4 <- train(classe ~ ., data=myTraining, method="lda")
predict4 <- predict(fit4, myTesting, type = "class")
confusionMatrix(predict4, myTesting$classe)$overall[1] #accuracy
```



