---
title: "Practical Machine Learning - Course Project"
author: "Rodrigo Rivera"
date: "October 3, 2016"
output: html_document
---

# Background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

#Data
The training data for this project are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv
The test data are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv
The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har.

##Load Libraries
```{r}
library(AppliedPredictiveModeling)
library(caret)
library(ElemStatLearn)
library(pgmm)
library(rpart)
library(gbm)
library(lubridate)
library(forecast)
library(e1071)
library(randomForest)
library(rattle)
```
##Download Data
```{r setup, include=FALSE}
trainUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
train <- read.csv(url(trainUrl), na.strings=c("NA","#DIV/0!",""))
testUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
test <- read.csv(url(testUrl), na.strings=c("NA","#DIV/0!",""))
dateDownloaded <- today()
```

#Cleaning Data
##Handle NAs
```{r}
colnames_train <- names(train)
colnames_test <- names(test)

nonNAs <- function(x) {
    as.vector(apply(x, 2, function(x) length(which(!is.na(x)))))
}

colcnts <- nonNAs(train)
drops <- c()
for (cnt in 1:length(colcnts)) {
    if (colcnts[cnt] < nrow(train)) {
        drops <- c(drops, colnames_train[cnt])
    }
}

train <- train[,!(names(train) %in% drops)]
train <- train[,8:length(colnames(train))]

test <- test[,!(names(test) %in% drops)]
test <- test[,8:length(colnames(test))]
```
##Finding Near Zervo Variance Variables
```{r}
nzvTrain <- nearZeroVar(train, saveMetrics=TRUE)
table(nzvTrain$zeroVar)
table(nzvTrain$nzv)
```
Since all are false, no variables are deleted.

##Creating Sets
The data already contains testing and triaining set, but I will make a new testing set, because the testing set's dims are much smaller than the training set.
```{r}
inTrain <- createDataPartition(y=train$classe, p=0.6, list=FALSE)
training <- train[inTrain,]
testing <- train[-inTrain,]
```


#Modeling
In this case the analysis will be done with two models: Classification Trees and Ranfom Forest
##Classification Trees
Model
```{r}
fit1 <- train(training$classe ~ ., data = training, method="rpart")
print(fit1, digits=3)
print(fit1$finalModel, digits=3)
fancyRpartPlot(fit1$finalModel)
```

Testing
```{r}
set.seed(500)
predict <- predict(fit1, newdata=testing)
print(confusionMatrix(predict, testing$classe)$overall[1], digits=4)
```
This model shows a very poor accuracy (49.22%). Will try this same model using crossed validation and preprocessing

Using Preprocess
```{r}
fit1pp <- train(training$classe ~ .,  preProcess=c("center", "scale"), data = training, method="rpart")
print(fit1pp, digits=3)
set.seed(500)
predict <- predict(fit1pp, newdata=testing)
print(confusionMatrix(predict, testing$classe)$overall[1], digits=4)
```
Same accuracy (49.22%)

Using Crossed Validation
```{r}
fit1cv <- train(training$classe ~ .,  trControl=trainControl(method = "cv", number = 4), data = training, method="rpart")
print(fit1cv, digits=3)
set.seed(500)
predict <- predict(fit1cv, newdata=testing)
print(confusionMatrix(predict, testing$classe)$overall[1], digits=4)
```

Using Both
```{r}
fit1b <- train(training$classe ~ .,  preProcess=c("center", "scale"), trControl=trainControl(method = "cv", number = 4), data = training, method="rpart")
print(fit1b, digits=3)
set.seed(500)
predict <- predict(fit1b, newdata=testing)
print(confusionMatrix(predict, testing$classe)$overall[1], digits=4)
```

##Random Forest
Model with Cross Validation
```{r}
fit2 <- train(training$classe ~ ., method="rf", trControl=trainControl(method = "cv", number = 4), data=training)
print(fit2, digits=3)
```

Testing
```{r}
set.seed(500)
predict2 <- predict(fit2, newdata=testing)
print(confusionMatrix(predict2, testing$classe)$overall[1], digits=4)
```

```{r}
fit2b <- train(training$classe ~ ., method="rf", preProcess=c("center", "scale"), trControl=trainControl(method = "cv", number = 4), data=training)
print(fit2b, digits=3)
predict2b <- predict(fit2, newdata=testing)
print(confusionMatrix(predict2b, testing$classe)$overall[1], digits=4)
```


#Testing The Best Model
The best model was clearly using Random Forest with Crossed Validation (although using both - Crossed Validation and Preprocess - had a very high accuracy), now I will test it with the provided test set to answer the quiz questions.
```{r}
prediction <- predict(fit2, newdata=test)
prediction
```

#Conclussion
The best model was random forest, with which I actually got 100% on the quiz.








